---
title: <center> P4 - Data Analysis </center>
author: <center> Ram Pangaluri, Ashok Shanker, Fabian Leon </center>
output: html_document
---


<!-- Changing the fonts and style to fit more into less pages -->
<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 18px;
  color: Black;
}
h1 { /* Header 1 */
  font-size: 18px;
  color: Black;
}
h2 { /* Header 2 */
    font-size: 16px;
  color: Black;
}
h3 { /* Header 3 */
  font-size: 14px;
  font-family: "Times New Roman", Times, serif;
  color: Black;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
h4.author {
  font-size: 12px;
  color: DarkBlue;
}
</style>

<!-- ## Grading Rubric -->
<!-- - Clear identification of at least 5 questions or story points that will be answered by your project  -->
<!-- - Clear statement of 4 or more different analyses of data that you will perform  -->
<!-- - Justification for each analysis, including why it is appropriate to the data and why it is appropriate to the story you are trying to tell.  -->
<!-- - Note: we expect at least one (probably more) of the analyses to be a visual analysis; that is,  understanding should be obtained through a visual representation (such as a plot or graph) of the data -->
<!-- - Presentation in an RMarkdown of 2 pages  -->

<!-- Begin document -->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## How have Texas' sorghum yields improved over time?
Advances in breeding, technology, production, and management all lead to higher sorghum yields. How have sorghum yields changed over time and what time periods saw the greatest increases in yield?

Range Sliders and Selectors in R will be implemented via line graphs. Plotting will be done with the plotly package. 



## What physiological dependencies are there between hybrid traits? 
For example, is there a relationship between a sorghum hybrid's height and how long it will take to flower?   


For example, is there a relationship between a sorghum hybrid's height and how long it will take to flower? An interactive scatterplot that plots and displays pearson’s R for whichever two variables the person selects. 

or 

Correlogram of hybrid traits  
Pearson's R correlation implemented in r with ggplot2 and ggpairs packages  


## What counties of Texas are the highest yielding environments for sorghum hybrids?  

A Chloropleth Map of average yields per county. The map would be dynamic and 

Required statistical tools:


## Grain Sorghum Traits in Texas Counties 





A nonlinear iterative partial least squares (nipals) principal component analysis (PCA) will be conducted on the calculated trait BLUPs attributable to counties from a linear mixed model analysis of Texas AgriLife data wherein traits of interest (yield, plant height, testweight, days to flowering, etc.) will be fitted to an all random model including year, county, hybrid, irrigation, 
 

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(equatiomatic)
library(lme4)

data = read.csv("data/data.csv")
str(data)

fit <- lm(GY ~ Hybrid + County + Hybrid*County + Total.Moisture + Previous.Crop + Maturity + Days.from.Plant.to.Harvest, data)
# show the theoretical model
equatiomatic::extract_eq(fit)

```




## Story point 5
Q:

Approach:

Required statistical tools:



